Semantic Linearity in a Lazy Optimising Compiler

### 2025-02-27 icfp.tex
preservation case ALT_
Finish type preservation proof typesetting

@@

>>>>

    \item[Case:] $\Gamma; \Delta \vdash (\Theta, \xl = y \mid x) \Downarrow (\Theta' \mid v) : \s, \Sigma$
    (1) $\G; \D \vdash \Theta, \xl = y \mid x : \s, \Sigma$\\
    (2) $\G, \Theta{\uparrow}^\omega; \D, \Theta{\uparrow}^1, \xl \vdash (x, y, \Theta{\downarrow}^1_e, \Sigma) : \s \otimes \s \bigotimes(\Theta{\downarrow^1_e}) \bigotimes(\Sigma)$\`by \ref{lem:unwrapenv}\\
    (3) $Let~\G';\D_1,\D_2,\D_3 = \G, \Theta{\uparrow}^\omega; \D, \Theta{\uparrow}^1$\` s.t. $(\D_2,\D_3)$ types $(\Theta{\downarrow}^1_e,\Sigma)$\\
    (4) $\G'; \xl \vdash x : \s$\\
    (5) $\G'; \D_1 \vdash y : \s$\\
    (6) $\G'; \D_1, \D_2, \D_3 \vdash (y, \Theta{\downarrow}^1_e, \Sigma) : \s \bigotimes(\Theta{\downarrow^1_e}) \bigotimes(\Sigma)$
    (7) $\G; \D \vdash \Theta \mid y : \s, \Sigma$\`by reconstructing\\
    (8) $\G; \D \vdash \Theta' \mid v : \s, \Sigma$\` by induction hypothesis\\
    \item[Case:] $\Gamma; \Delta \vdash (\Theta, x{:}_{\D'}\s = e \mid x) \Downarrow (\Theta', x{:}_{\D_a}\s = v \mid v) : \s, \Sigma$
    (1) $\Gamma; \Delta \vdash (\Theta, x{:}_{\D'}\s = e \mid x) : \s, \Sigma$\\
    (2) $\G, \Theta{\uparrow}^\omega; \D, \Theta{\uparrow}^1 \vdash \llet{x{:}_{\D'} = e}{(x, \Theta{\downarrow}^1_e, \Sigma)} : \s \bigotimes(\Theta{\downarrow}^1_\s) \bigotimes(\Sigma)$\`by \ref{lem:unwrapenv}\\
    (3) $Let~\D',\D'' = \D,\Theta{\uparrow}^1$\\
    (4) $\G, \Theta{\uparrow}^\omega; \D' \vdash e : \s$\`by inv on let (2)\\
    (5) $\G, \Theta{\uparrow}^\omega; \D', \D'' \vdash (x, \Theta{\downarrow}^1_e, \Sigma) : \s \bigotimes(\Theta{\downarrow}^1_\s) \bigotimes(\Sigma)$\`by inv on let (2)\\
    (6) $\G, \Theta{\uparrow}^\omega; \D' \vdash x : \s$\\
    (7) $\G, \Theta{\uparrow}^\omega; \D'' \vdash (\Theta{\downarrow}^1_e, \Sigma) : \bigotimes(\Theta{\downarrow}^1_\s) \bigotimes(\Sigma)$\\
    (8) $\G, \Theta{\uparrow}^\omega; \D', \D'' \vdash (e, \Theta{\downarrow}^1_e, \Sigma) : \tau \bigotimes(\Theta{\downarrow}^1_\s) \bigotimes(\Sigma)$\` by pair (4,7)\\
    (9) $\G; \D \vdash \Theta \mid e : \tau,\Sigma$\` by reconstructing\\
    (10) $\G; \D \vdash \Theta \mid v : \tau,\Sigma$\` by induction hypothesis\\
    (11) $Let~\G';\D_a,\D_b = \G, \Theta'{\uparrow}^\omega, \D, \Theta'{\uparrow}^1$\\
    (11a) $\G'; \D_a \vdash v : \s$
    (11b) $\G'; \D_b \vdash (\Theta'{\downarrow}^1_e, \Sigma) : \bigotimes(\Theta'{\downarrow}^1_\s) \bigotimes(\Sigma)$\\
    (12) $\G'; \D_a \vdash \llet{x{:}_{\D_a}\s = v}{v} : \s$\`by let (11a,11a)\\
    (13) $\G; \D \vdash \Theta' \mid \llet{x{:}_{\D_a}\s = v}{v} : \s, \Sigma$\`by reconstructing\\
    (14) $\G; \D \vdash \Theta', x{:}_{\D_a}\s = v \mid v : \s, \Sigma$\`by env. expansion\\
Note: It's slightly more elaborate to accommodate recursion. The runtime
environment bindings must be tagged with the recursive group such that the
environment expansion can decide whether to unwrap to a letrec group or
not. Unlike the case proved here, when the $\D$-binding belongs to a let group,
the whole group is unwrapped and we can invert the assumption to get an
$e$ typed with the $x$s from the group.\\

@@

\item[Case:] $\Gamma; \Delta \vdash (\Theta \mid \ccase{e}{z{:}_{\Delta'}\s~\{\_ \to e'}\}) \Downarrow (\Theta'' \mid v) : \tau$
    (1) $\Gamma; \Delta \vdash \Theta \mid \ccase{e}{z{:}_{\Delta_a}\s~\{\_ \to e'}\} : \tau, \Sigma$\\
    (2) $\G, \Theta{\uparrow}^\omega; \D, \Theta{\uparrow}^1 \vdash (\ccase{e}{z{:}_{\Delta_a}\s~\{\_ \to e'}\}, \Theta{\downarrow}^1_e, \Sigma) : \tau \otimes \bigotimes(\Theta{\downarrow}^1_\sigma) \otimes \bigotimes(\Sigma)$\`by \ref{lem:unwrapenv}\\
    (4) $\G'; \D_a, \D_b \vdash \ccase{e}{z{:}_{\D_a}\s\{\_ \to e'\}} : \tau$\\
    (5) $\G'; \D_a \vdash e : \sigma$\\
    Subcase $e$ is in WHNF\\
    (-) Similar to the same subcase in the case-with-pattern-vars reduction,\\
        except only $Alt_{\_}$ is relevant which is far simpler\\
    (6) $\G', z{:}_{\irr{\D_a}}\s; \irr{\D_a}, \D_b \vdash e' : \tau$\\
    (7) $\G'; \D_b, z{:}_1\s \vdash e' : \tau$\` by $\D$-bound to linear\\
    (8) $\G'; \D_a, \D_b, z{:}_1\s \vdash (e, e') : \sigma \otimes \tau$\` by pair\\
    (9) $\G; \D, z{:}_1\s \vdash \Theta \mid e : \sigma, e' : \tau, \Sigma$\` by reconstructing\\
    (10) $\G; \D, z{:}_1\s \vdash \Theta' \mid K~\ov{x_i} : \sigma, e' : \tau, \Sigma$\` by induction hypothesis\\
    (11) $Let~\G';\D_c,\D_d,\D_4,\D_5 = \G, \Theta'{\uparrow}^\omega; \Delta, \Theta'{\uparrow}^1$\`s.t. $(\D_4,\D_5)$ types $(\Theta'{\downarrow}^1_e, \Sigma)$\\
    (12) $\G';\D_c \vdash K~\ov{x_i} : \s$\\
    (13) $\G';\D_d,z{:}_1\s \vdash e' : \tau$\\
    (14) $\G';\D_c, \D_d \vdash e'[K~\ov{x_i}/z] : \tau$\` by $1$-subst lemma\\
    (15) $\G;\D \vdash \Theta' \mid e'[K~\ov{x_i}/z] : \tau, \Sigma$\` by reconstructing\\
    (16) $\G;\D \vdash \Theta'' \mid v : \tau, \Sigma$\` by induction hypothesis\\
